---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sdmxdata

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/sdmxdata)](https://CRAN.R-project.org/package=sdmxdata)
[![R-CMD-check](https://github.com/edwindj/sdmxdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/edwindj/sdmxdata/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Introduction

_Work in progress, not to be used for production work_

Why `sdmxdata`? The package is designed to provide a simple interface to
retrieve and select data from open statistical data providers that support the SDMX v2.1 API. 

It focusses on:

- retrieving lists of dataflows: `list_dataflows`, for getting an overview of the
available tables (dataflow) at a data provider, including their descriptions.

- retrieving the structure of a dataflow: `get_dataflow_structure`, for getting
detailed information the structure and metadata of a dataflow.

- retrieving data: `get_data` and `get_observations` 
for selecting and retrieving data from a dataflow. Both functions allow for 
easy filtering on dimensions, reporting periods, and other metadata.
 
- retrieving lists of agencies and subagencies: `list_agencies`, for getting an overview of the
agencies 

It does not implement a full fledged SDMX metadata model or data model. For that, you can use [`rsdmx`](https://CRAN.R-project.org/package=rsdmx).

## Installation

You can install the development version of sdmxdata from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("edwindj/sdmxdata")
```

## Example


```{r setup}
library(sdmxdata)
```

To retrieve a list of tables:

```{r}
CBSbeta <- sdmxdata::get_endpoint("NL1")
dfs <- CBSbeta |> list_dataflows()
dfs[,1:5] |> 
  head()
```
```{r}
OECD <- sdmxdata::get_endpoint("OECD")
dfs_oecd <- OECD |> list_dataflows()
dfs_oecd[,1:5] |> 
  head()
```
To retrieve a dataflow structure:

```{r}
dataflow_ref <- dfs$ref[4]
print(dataflow_ref)

dsd <- CBSbeta |> get_dataflow_structure(ref = dataflow_ref)
dsd
```

To retrieve data use `get_data`

```{r}
data <- CBSbeta |> get_data(agencyID=dsd$agencyID, id = dsd$id, pivot="Topics")
head(data)
```

Or get the underlying observations with `get_observations`

```{r}
obs <- CBSbeta |> get_observations(id="DF_37230ned", agencyID="NL1.CNVT")
```

```{r}
head(obs)
```

To retrieve a list of agencies or subagencies:

```{r}
agencies <- CBSbeta |> list_agencies()
head(agencies)
```

